### 2021-10-13 publication/abstract.tex
scaffold publication

\section*{Technical Abstract}
Labelled networks form a very common and important class of data,
naturally appearing in numerous applications in science and engineering.
A typical inference goal is to determine how the vertex labels
(or {\em features}) affect the network's structure.

@@

Future work on this topic could benefit from extending the FFBM to be hierarchical in nature. Such a modification would be relatively natural and would increase the descriptive power of the model. The current FFBM formulation is limited in that it can only explain structure at the macro-scale. Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.

>>>>

\begin{abstract}
Labelled networks form a very common and important class of data,
naturally appearing in numerous applications in science and engineering.
A typical inference goal is to determine how the vertex labels
(or {\em features}) affect the network's structure.

@@

Future work on this topic could benefit from extending the FFBM to be hierarchical in nature. 
Such a modification would be relatively natural and would increase the descriptive power of the model. 
The current FFBM formulation is limited in that it can only explain structure at the macro-scale. 
Nevertheless, a hierarchical formulation would allow explanations of structure at any arbitrary length-scale.
\end{abstract}

### 2021-10-13 publication/ffbm.tex
scaffold publication

Let $N$ denote the number of vertices, $B$ the number of blocks in the graph,
and ${\cal X}$ the set of values each feature can take.
We define the vector $x_i \in \Xcal^D$ as the feature vector of vertex $i$, 
where $D$ is the number of features associated with each vertex.
For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
the feature vectors $\{x_i\}_{i=1}^{N}$ 
as its rows.

@@

Once the block memberships $b$ have been generated, we then draw the 
graph $A$ from the microcanonical DC-SBM with additional parameters 
$\psi = \{\psi_e, \psi_k\}$:
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
	\label{eqn:A-generation}
\end{equation}

>>>>

Let $N$ denote the number of vertices, $B$ the number of blocks in the graph,
and ${\cal X}$ the set of values each feature can take.
We define the vector $x_i \in \Xcal^D$ as the feature vector for vertex $i$, 
where $D$ is the number of features associated with each vertex.
For the datasets we analyse, we deal with binary feature flags so $\Xcal = \{0, 1\}$. We write $X$ for the $N\times D$ {\em feature matrix} containing
the feature vectors $\{x_i\}_{i=1}^{N}$ 
as its rows.

@@

Once the block memberships $b$ have been generated, we then draw the 
graph $A$ from the microcanonical DC-SBM with additional parameters 
$\psi = \{\psi_e, \psi_k\}$:
%
\begin{equation}
	A \sim \textrm{DC-SBM}_{\textrm{MC}} (b, \psi_e, \psi_k).
	\label{eqn:A-generation}
\end{equation}
%
Figure~\ref{fig:ffbm} illustrates the proposed model.

### 2021-10-13 publication/introduction.tex
scaffold publication

To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks. Nevertheless, this approach is limited in that it can only consider one feature at a time. This makes it difficult to rank-order the features by  magnitude of impact. Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit. We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.

Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM). This is an extension of the SBM to labelled networks. We go on to present an efficient algorithm for sampling from the parameters of the feature-to-block generator, and to describe how the 
sampled FFBM parameters can be interpreted to determine which features have the largest impact on overall graphical structure.

>>>>

To analyse a labelled network using one of the simple SBM variants, a typical inference procedure would be to first partition the graph into blocks grouped by distinct values of the feature of interest, and then use the associated model to test for evidence of heterogeneous connectivity between the feature-grouped blocks. Nevertheless, this approach is limited as it can only consider one feature at a time. This makes it difficult to rank-order the features by  magnitude of impact. Lastly, the feature-grouped blocks are often an unnatural partition of the graph, leading to a poor model fit. We would instead prefer to partition the graph into its most natural blocks and then find which of the available features -- if any -- best predict the resulting partition.

Thus motivated, we present a novel framework for modelling labelled networks, which we call the feature-first block model (FFBM). This is an extension of the SBM to labelled networks. We go on to present an efficient algorithm for sampling from the FFBM parameters, and to describe how the sampled parameters can be interpreted to determine which features have the largest impact on overall graphical structure.
